Motmot: A Distributed Chat Protocol

### 2012-05-04 paper/paper.tex
changed a word
libmotmot testing

@@

We were unable to fully integrate the \libmotmot protocol in the plugin.
The code is outlined with in the plugin file, however it is not currently
functional. Hopefully this will be fixed in future versions so that
\libmotmot can be used as a \libpurple plugin.

>>>>

Testing \libmotmot in a realistic setting was also challenging. We resorted to
client ``fuzzing'' as our primary testing strategy, in which we randomly issued
valid commands to a pool of connected \Motmot instances. These commands included
messages, join invitations, and parts, and were issued at random intervals to
random chat participants.

One of the biggest challenges was writing a testing framework that allowed us to
script interactions between many processes. Our first attempt was a Ruby DSL
that allowed small scripts to spawn and interact with a given number of \Motmot
instances. After performing arbitrary manipulations to the client pool,
including sending messages, creating new clients, and killing clients, a reaper
thread ensures message ordering.

The second iteration of a testing framework instead relied more heavily on UNIX
abstractions. First, we create a UNIX named pipe to act as \Motmot's input
channel (replacing \texttt{stdin}). A small utility program
(\texttt{test/motmot}) then assists in \texttt{dup2}-ing the correct file
descriptors in place and spawning the correct process. This allows many
\Motmot's to be spawned in separate windows, while centralizing the control of
the instances. Unlike the previous testing framework, \texttt{stdout} is not
checked. Finally, a central trace runner writes commands to the appropriate
named pipes, causing messages to be sent to various windows.

A key debugging feature of both of these frameworks was the ability to attach
debuggers to every \Motmot instance spawned. This served to be a critical
debugging tool, especially when analyzing segmentation faults or ensuring client
state consistency.

One final problem when testing \libmotmot was in simulating the network.  Chats
occur over TCP in the real world, which have some amount of latency and can
perform packet fragmentation, neither of which is a concern when interfacing
over UNIX domain sockets. To simulate a lossy network (on which connections
occasionally die), we wrote a network proxy, which ferried data from one UNIX
socket to another, fragmenting them and adding between 5 and 500 milliseconds of
latency. Connections were also occasionally dropped at some random interval.

While simulating a network proved difficult, we found that fuzzing with network
delays was a very effective way of finding bugs. After sending tens of millions
of messages through \libmotmot, we are fairly confident that it correctly
implements Paxos and delivers chats reliably, even at very high message
throughputs (on the order of a thousand messages a second on low-latency
networks).

@@

We were unable to fully integrate the \libmotmot protocol in the plugin.
The code is outlined with in the plugin file, however it is not currently
functional. Hopefully this will be fixed in future versions so that
\libmotmot can be used within a \libpurple plugin.
